# Characteristics of the tree plantations

We assessed whether the native (*Araucaria angustifolia*) and exotic (*Pinus elliottii*) tree plantations differ in three aspects: age, understorey vegetation obstruction, and distance to the nearest natural forest patch larger than 10 hectares.

### Ensuring reproducibility {.unnumbered}
```{r}
set.seed(26)
```

### Loading packages {.unnumbered}
```{r, output = FALSE}
library(stats)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(rstatix)
```

### Importing data {.unnumbered}
```{r}
#| code-overflow: wrap

# https://knb.ecoinformatics.org/view/doi:10.5063/F1QR4VK1
setwd("C:/Users/ivana/OneDrive/FLONA_aves/data")
data <- read.csv("FLONA-PF_dataset.txt", sep = "\t")
env_all <- data[,1:8]
env_plant <- subset(env_all, forest.type != "Natural_forest")
```

### Preparing data {.unnumbered}
```{r}
color <- rep(c("#912DBF","#BF912D"), each = 6)
color_complete <-rep(c("#2DBF91","#912DBF","#BF912D"), each = 6)
```

## Age
### Performing a t-test
```{r}
#| code-overflow: wrap
# Perform the Shapiro-Wilk test to check if age data is normally distributed
shapiro.test(env_plant$age) # It's normal (p > 0.05)

# Perform Levene's test to check for homogeneity of variances across forest types
levene_test(age ~ forest.type, data = env_plant) # The variances are homogeneous (p > 0.05)

# Perform a t-test to compare age between forest types
t.test(env_plant$age ~ env_plant$forest.type, var.equal = TRUE)
```
There was a significant difference in age between the two plantation types (t = 4.6696, p < 0.001).

### Plotting the graph
```{r}
#| code-overflow: wrap
plot_age <-
  ggplot(data = env_plant,
         aes(x = forest.type, y = age)) +
  
  scale_y_continuous(breaks = c(50, 55, 60, 65, 70, 75)) + 
  
  geom_boxplot(outlier.shape = NA,
               width = 0.7, col = "black", fill = c("#912DBF","#BF912D"), 
               alpha = 0.30, show.legend = FALSE) + 
  
  geom_jitter(shape= 21, 
              width = 0.2, height = 0.2, colour = "black", 
              fill = color, size = 2.5) + 
  
  scale_x_discrete(labels = c("Araucaria\nplantation",
                              "Pine\nplantation")) +
  
  labs(x = " ", y = "Age in 2021 (years)", colour="black",
       tag = "a") + 
  
  theme_pubr (base_size = 10) + 
  
  theme(axis.line = element_line(linewidth = 1/3))+
  
  theme(plot.tag = element_text(size = 16, face = "bold"))

plot_age
```

## Distance to forest
### Performing a t-test
```{r}
#| code-overflow: wrap
# Perform the Shapiro-Wilk test to check if the distances to the nearest natural forest are normally distributed
shapiro.test(env_plant$distance.forest.m) # It's normal (p > 0.05)

# Perform Levene's test to check for homogeneity of variances across forest types
levene_test(distance.forest.m ~ forest.type, data = env_plant) # The variances are homogeneous (p > 0.05)

# Perform a t-test to compare the distances to the nearest natural forest between forest types
t.test(env_plant$distance.forest.m ~ env_plant$forest.type, var.equal = TRUE)
```
The distance of the tree plantations to the nearest natural forest did not differ significantly between the two types of plantation (t = -0.121; p = 0.906). 

### Plotting the graph
```{r}
#| code-overflow: wrap
plot_dist <- 
  
  ggplot(data = env_plant,
         aes(x = forest.type, y = distance.forest.m)) +
  
  geom_boxplot(outlier.shape = NA,
               width = 0.7, col = "black", 
               fill = c("#912DBF", "#BF912D"),
               alpha = 0.30, show.legend = FALSE) +
  
  geom_jitter(shape= 21, width = 0.2,
              height = 0.2, colour = "black", 
              fill = color, size = 2.5) +
  
  scale_x_discrete(labels = c("Araucaria\nplantation",
                              "Pine\nplantation")) +
  
  scale_y_continuous(breaks = c(200, 600, 1000, 1400, 1800)) +
  
  labs(x = " ", 
       y = "Distance to natural forest (m)", 
       colour="black",
       tag = "b") +
  
  theme_pubr (base_size = 10) + 
  
  theme(axis.line = element_line(linewidth = 1/3)) +
  
  theme(plot.tag = element_text(size = 16, face = "bold"))

plot_dist
```

## Understory vegetation obstruction
### Fitting an ANOVA model
```{r}
#| code-overflow: wrap
# Perform the Shapiro-Wilk test to check if understorey vegetation obstruction is normally distributed
shapiro.test(env_all$obstruction) # It's normal (p > 0.05)

# Perform Levene's test to check for homogeneity of variances across forest types
levene_test(obstruction ~ as.factor(forest.type), data = env_all) # The variances are homogeneous (p > 0.05)

# Perform one-way ANOVA to test for differences in understorey vegetation obstruction between forest types
anova <- aov(obstruction ~ forest.type, data = env_all)
summary(anova)

# Perform the Shapiro-Wilk test on the residuals to check if they are normally distributed
shapiro.test(anova$residuals) # Residuals of the ANOVA model are normal (p > 0.05)

# Perform Levene's test on residuals to check for homogeneity of variances
levene_test(anova$residuals ~ as.factor(forest.type), data = env_all) # The variances are homogeneous (p > 0.05)

# Perform Tukey's HSD post-hoc test to compare pairwise differences between forest types
par_par <- TukeyHSD(anova)
par_par
```

Understorey vegetation obstruction significantly differed between forest types (F = 5.986, p = 0.01). We found significant differences between pine plantations and natural forests (diff = -13.50, p = 0.015) and between pine plantations and araucaria plantations (diff = -11.33, p = 0.041). No significant difference was found between araucaria plantations and natural forests (diff = 2.16, p = 0.864).

### Plotting the graph
```{r}
#| code-overflow: wrap
# Renaming forest types for consistent ordering in the plot
env_all$forest.type[env_all$forest.type == "Natural_forest"] <- "1.Natural_forest"

env_all$forest.type[env_all$forest.type == "Araucaria_plantation"] <- "2.Araucaria_plantation"

env_all$forest.type[env_all$forest.type == "Pine_plantation"] <- "3.Pine_plantation"

plot_obs <- 
  ggplot(data = env_all,
         aes(x = forest.type, y = obstruction)) +
  
  geom_boxplot(outlier.shape = NA,
               width = 0.7, col = "black", 
               fill = c("#2DBF91", "#912DBF", "#BF912D"), 
               alpha = 0.30, show.legend = FALSE) +
  
  geom_jitter(shape= 21, 
              width = 0.2, height = 0.2, colour = "black", 
              fill = color_complete, size = 2.5) +
  
  scale_x_discrete(labels = c("Natural \nforest",
                              "Araucaria\nplantation",
                              "Pine\nplantation")) +
  
  labs(x = " ", y = "Understorey obstruction (%)", colour="black",
       tag = "c") + 
  
  theme_pubr (base_size = 10) + 
  
  theme(axis.line = element_line(linewidth = 1/3)) +
  
  theme(plot.tag = element_text(size = 16, face = "bold"))

plot_obs

```

### Saving the graph
```{r}
#| code-overflow: wrap
ggsave(
  plot = plot_age + plot_dist + plot_obs,
  filename = "charact_plantation.png", dpi = 500,
  width = 7 * 3, height = 7, units = 'cm')
```