# Bird Species Abundance

Predictor variables:

1. **Forest type** - We used the Kruskal-Wallis test to assess the effect of forest type on species abundances. We then applied the pairwise Wilcoxon test to identify significant differences between forest types. 

2. **Understorey vegetation obstruction** - We performed simple linear regression analyses to investigate the effect of understorey vegetation obstruction on species abundance.

### Ensuring Reproducibility {.unnumbered}
```{r}
set.seed(26)
```

### Load packages {.unnumbered}
```{r, output = FALSE}
library(ggplot2)
library(ggpubr)
library(dplyr)
```

### Importing data {.unnumbered}
```{r, output = FALSE}
# https://knb.ecoinformatics.org/view/doi:10.5063/F1QR4VK1
setwd("C:/Users/ivana/OneDrive/FLONA_aves/data")
data <- read.csv("FLONA-PF_dataset.txt", sep = "\t")
abund <- data[,c(1:2, 9:59)]
env_all <- data[,1:8]
```

### Preparing data {.unnumbered}
```{r}
# Renaming forest types for consistent ordering in the plot
abund$forest.type <- recode(abund$forest.type,
                           "Natural_forest" = "1.Natural_forest",
                           "Araucaria_plantation" = "2.Araucaria_plantation",
                           "Pine_plantation" = "3.Pine_plantation")
```

## Forest type

### Preparing data
```{r, output = FALSE}
#| code-overflow: wrap
# Calculates species abundance per sample
abund["abundance"] <- rowSums(data[9:59])
```

### Calculating ANOVA
```{r, output = FALSE}
#| code-overflow: wrap
# Performs one-way ANOVA to test for differences in species abundance among forest types
anova <- aov(abundance ~ forest.type, data = abund)
```

### Testing normality of ANOVA residuals
```{r}
#| code-overflow: wrap
# Tests whether ANOVA residuals are normally distributed
shapiro.test(anova$residuals) # It's not normal (p < 0.05)
```

Since the ANOVA residuals were not normally distributed, we used the Kruskal–Wallis test instead. To compare groups, we then ran pairwise Wilcoxon tests.

### Calculating Kruskal-Wallis test
```{r}
#| code-overflow: wrap
# Performs the Kruskal-Wallis test to compare species abundance among forest types
kruskal <- kruskal.test(abundance ~ forest.type, data = abund)
kruskal
```

The number of individuals varied significantly between forest types (Kruskal-Wallis test: χ² = 10.561, df = 2, p = 0.005).

### Calculating Wilcoxon test
```{r}
#| code-overflow: wrap
# Performs pairwise Wilcoxon tests to compare forest types, with p-value adjustment using the Benjamini-Hochberg method
pair <- pairwise.wilcox.test(abund$abundance, abund$forest.type, p.adjust.method = "BH")
pair
```

Wilcoxon pairwise comparisons showed that species were more abundant in natural forests than in araucaria plantations (p = 0.023) and pine plantations (p = 0.023). There was also a difference in species abundance between the two plantation types (p = 0.041).


### Plotting the graph
```{r}
#| code-overflow: wrap
# Creates a vector of colours, assigning one colour to each forest type (6 samples per type)
color <- rep(c("#2DBF91", "#912DBF", "#BF912D"), each = 6)

plot_abund1 <- 
  ggplot(data = abund,
         aes(x = forest.type, y = abundance)) +
  
  geom_boxplot(outlier.shape = NA,
               width = 0.7, col = "black", 
               fill = c("#2DBF91", "#912DBF","#BF912D"),
               alpha = 0.30, show.legend = FALSE) +
  
  geom_jitter(shape = 21, width = 0.2, 
              height = 0.2, colour = "black", 
              fill = color, size = 2.5) +
  
  scale_x_discrete(labels = c("Natural\nforest", 
                              "Araucaria\nplantation",
                              "Pine\nplantation"))+
  
  labs(x = " ", y = "Number of individuals", 
       colour="black", 
       tag = "b") +
  
  theme_pubr (base_size = 10) + 
  
  theme(axis.line = element_line(linewidth = 1/3)) +
  
  theme(plot.tag = element_text(size = 16, face = "bold")) 
```


### Saving the graph
```{r}
plot_abund1
ggsave(
  plot = plot_abund1,
  filename = "abundance_foresttype.png", dpi = 500,
  width = 7, height = 7, units = 'cm')
```


## Understory vegetation obstruction

### Preparing data
```{r}
#| code-overflow: wrap

# Creates a new data frame with forest type, species abundance, and understorey vegetation obstruction
set <- data.frame(forest.type = env_all$forest.type,
                  abundance = abund$abundance,
                  obstruction = env_all$obstruction)
```

### Fitting the regression model
```{r}
#| code-overflow: wrap
# Fits a linear model to assess the effect of vegetation obstruction on species abundance
reg_abund <- lm(set$abundance ~ set$obstruction)
summary(reg_abund)

# Tests the normality of model residuals and plots their histogram
shapiro.test(reg_abund$residuals)
```

### Plotting the graph
```{r}
#| code-overflow: wrap
plot_abund2 <-
  ggplot(set, 
         aes(x = obstruction, y = abundance,
             fill = forest.type)) +
  
  scale_y_continuous(limits = c(NA, 100),
                     breaks = c(20, 40, 60, 80, 100)) +
  
  geom_smooth(method = "lm", formula = y ~ x, 
              se = TRUE, color = "black", 
              fill = "gray", fullrange = TRUE) +
  
  geom_point(colour = "black", size = 2.5, shape = 21) +
  
  scale_fill_manual(values = c("#812DBF", "#2DBF81", "#BF812D"), 
                    labels = c("Araucaria plantation", "Natural forest",
                               "Pine plantation")) +
  
  labs(x = "Understorey vegetation obstruction (%)", 
       y = "Number of individuals",
       tag = "b") +
  
  theme_pubr(base_size = 10) +
  
  theme(legend.position = "none") +
  
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

### Saving the graph
```{r}
#| code-overflow: wrap
plot_abund2 
ggsave(
  plot = plot_abund2,
  filename = "abundance_understory.png", dpi = 500,
  width = 7, height = 7, units = 'cm')
```


## Indicator species

### Preparing data
```{r}
# Selects only indicator species
abund <- abund[,c(1,6,21,38,48,53)]

forest <- abund[1:6, -1]
forest <- colSums(forest)
forest <- as.data.frame(forest)
colnames(forest)[1] <- "forest.type"


araucaria <- abund[7:12, -1]
araucaria <- colSums(araucaria)
araucaria <- as.data.frame(araucaria)
colnames(araucaria)[1] <- "forest.type"

pinus <- abund[13:18, -1]
pinus <- colSums(pinus)
pinus <- as.data.frame(pinus)
colnames(pinus)[1] <- "forest.type"

# Creates new data.table
forest.type <- rbind(forest, araucaria, pinus)
colnames(forest.type)[1] <- "abundance"

forest.type$forest.type <- NA
forest.type$forest.type[1:5] <- "3_natural_forest"
forest.type$forest.type[6:10] <- "2_araucaria_plantation"
forest.type$forest.type[11:15] <- "1_pine_plantation"

forest.type$species <- rownames(forest.type)
forest.type <- forest.type[,c(3,1,2)]

rownames(forest.type) <- NULL

forest.type$species[c(1,6,11)] <- "4_Chiroxiphia_caudata"
forest.type$species[c(2,7,12)] <- "5_Leptopogon_amaurocephalus"
forest.type$species[c(3,8,13)] <- "3_Synallaxis_cinerascens"
forest.type$species[c(4,9,14)] <- "1_Turdus_leucomelas"
forest.type$species[c(5,10,15)] <- "2_Zonotrichia_capensis"
```

### Plotting the graph
```{r}
#| code-overflow: wrap
plot_indsp <-
  ggplot(forest.type, 
         aes(x = species, y = abundance, fill = forest.type)) +
  
  scale_fill_manual(values = c("3_natural_forest" = "#2DBF81",
                               "2_araucaria_plantation" = "#812DBF",
                               "1_pine_plantation" = "#BF812D"),
                    labels = c("1_pine_plantation" = "Pine plantation",
                               "2_araucaria_plantation" = "Araucaria plantation",
                               "3_natural_forest" = "Natural forest")) +
  
  geom_bar(stat = "identity", position = "fill",
           alpha = 0.30,
           colour = "black",
           width = 0.7) +
  
  coord_flip() +
  
  scale_y_continuous(labels = scales::percent) +
  
  scale_x_discrete(labels = c("5_Leptopogon_amaurocephalus" = "Leptopogon amaurocephalus",
                              "4_Chiroxiphia_caudata" = "Chiroxiphia caudata",
                              "3_Synallaxis_cinerascens" = "Synallaxis cinerascens",
                              "2_Zonotrichia_capensis" = "Zonotrichia capensis",
                              "1_Turdus_leucomelas" = "Turdus leucomelas")) +
  
  labs(x = " ", y = "Individuals captured by forest type", fill = " ") +
  
  theme_pubr(base_size = 12) +
  
  theme(panel.border = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom",
        axis.text.y = element_text(face = "italic")) +
  
  guides(fill = guide_legend(reverse = TRUE))
```

### Saving the graph
```{r}
plot_indsp
ggsave(
  plot = plot_indsp,
  filename = "abund_sp_indicadoras.png", dpi = 500,
  width = 7*3, height = 7, units = 'cm')
```